### 2011-05-23 tags/crisp-1.1.6/trunk/Papers/experiences/experiments.tex
[maven-release-plugin]  copy for tag crisp-1.1.6

In the first experiment, we generate a series of sentence generation
problems which require the planner to compute a plan representing the
sentence ``Mary likes the Adj$_1$ \ldots Adj$_n$ rabbit.''  Each problem
instance assumes a certain number $m$ of rabbits that were distinguished by
$n \leq m$ different properties, such that all $n$ properties are required
to distinguish the target referent from all other rabbits.  The $n$
properties are realized as $n$ different adjectives, in any order.  This
setup allows us to control the plan length (a plan with $n$ properties will
have length $n+4$) and the universe size (the universe will contain $m+1$
individuals in addition to the differently-typed individuals used to encode
the grammar).

@@

In fact, the time spent by FF on grounding is such that it is consistently
outperformed by an ad-hoc Java implementation of GraphPlan which only
computes instances of predicates and actions as they are discovered (i.e.,
while the planning graph is being built)---despite the fact that FF is
consistently much faster as far as pure search time is concerned.  Looking
at the difference from another angle, FF's performance is much more
sensitive to the domain size: if we fix $n=1$, FF takes 60 ms to compute a
plan at $m=1$, but 2.4 seconds (for the same plan) at $m=10$; our GraphPlan
implementation takes 30 ms at $m=1$ and still only requires 50 ms at
$m=10$.  Conversely, GraphPlan's runtime grows much faster with the plan
size (i.e., with growing values of $n$ for a fixed $m$).  Larger (but still
realistically-sized) instances of the sentence generation problem are still
problematic for the planners we tested.

@@

Results for the $h=20$ case, with $n$ ranging from $1$ to $40$, are
shown in Figure~\ref{fig:give-runtime-minimal}.  The most obvious
result is that FF is unable to solve any problems beyond $n=13$ on our
experimentation machine within the memory limit of 1 GB.  SGPLAN, on
the other hand, solves instances beyond $n=40$ without major
problems.\footnote{Our ad-hoc implementation of GraphPlan is unable to
  find any of these plans.}  The time spent on grounding is not a
major factor in either planner, probably because the planners need
more time to actually compute the plan---for instance, the optimal
plan for the problem $n=40$ has a length of about 1600 steps.  As an
example, the plan for $n=h=1$ looks as follows:

\begin{enumerate}
\item $\mathsf{move}(\mathsf{pos\_1\_1},\mathsf{pos\_1\_2}, \mathsf{north})$
\item $\mathsf{manipulate}\textsf{-}\mathsf{u1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_1\_2})$
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{north}, \mathsf{east})$
\item $\mathsf{move}(\mathsf{pos\_1\_2}, \mathsf{pos\_2\_2},
  \mathsf{east})$
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{east}, \mathsf{south})$
\item $\mathsf{move}(\mathsf{pos\_2\_2}, \mathsf{pos\_2\_1}, \mathsf{south})$
\item $\mathsf{manipulate}\textsf{-}\mathsf{l1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_2\_1})$
\end{enumerate}

>>>>

In the first experiment, we construct a series of sentence generation
problems which require the planner to compute a plan representing the
sentence ``Mary likes the Adj$_1$ \ldots Adj$_n$ rabbit.''  Each problem
instance assumes a certain number $m$ of rabbits that were distinguished by
$n \leq m$ different properties, such that all $n$ properties are required
to distinguish the target referent from all other rabbits.  The $n$
properties are realized as $n$ different adjectives, in any order.  This
setup allows us to control the plan length (a plan with $n$ properties will
have length $n+4$) and the universe size (the universe will contain $m+1$
individuals in addition to the differently-typed individuals used to encode
the grammar).

@@

In our examples, the time spent by FF on grounding is such that it is
consistently outperformed by an ad-hoc Java implementation of GraphPlan
which only computes instances of predicates and actions as they are
discovered (i.e., while the planning graph is being built)---despite the
fact that FF is consistently much faster as far as pure search time is
concerned.  Looking at this difference from another angle, FF's performance
is much more sensitive to the domain size: if we fix $n=1$, FF takes 60 ms
to compute a plan at $m=1$, but 2.4 seconds to compute the same plan at
$m=10$; our GraphPlan implementation takes 30 ms at $m=1$ and still only
requires 50 ms at $m=10$.  Conversely, GraphPlan's runtime grows much
faster with the plan size (i.e., with growing values of $n$ for a fixed
$m$).  Larger, but still realistically-sized, instances of the sentence
generation problem are still problematic for the planners we tested.

@@

Results for the $h=20$ case, with $n$ ranging from $1$ to $40$, are
shown in Figure~\ref{fig:give-runtime-minimal}.  The most obvious
result is that FF is unable to solve any problems beyond $n=13$ on our
experimentation machine within the memory limit of 1 GB.  SGPLAN, on
the other hand, solves instances beyond $n=40$ without major
problems.\footnote{Our ad-hoc implementation of GraphPlan is unable to
  find any of these plans.}  The time spent on grounding is not a
major factor in either planner, probably because the planners need
more time to actually compute the plan---for instance, the optimal
plan for the problem $n=40$ has a length of about 1600 steps.  As a
concrete example, the following is a plan for $n=h=1$:

\begin{enumerate}
\item $\mathsf{move}(\mathsf{pos\_1\_1},\mathsf{pos\_1\_2}, \mathsf{north})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{u1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_1\_2})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{north}, \mathsf{east})$,
\item $\mathsf{move}(\mathsf{pos\_1\_2}, \mathsf{pos\_2\_2},
  \mathsf{east})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{east}, \mathsf{south})$,
\item $\mathsf{move}(\mathsf{pos\_2\_2}, \mathsf{pos\_2\_1}, \mathsf{south})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{l1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_2\_1})$.
\end{enumerate}

### 2011-05-23 tags/crisp-1.1.6/trunk/Papers/operator-splitting/opsplit-definition-v2.tex
[maven-release-plugin]  copy for tag crisp-1.1.6

For example, consider the blocksworld operator ``move(A,B,C)'', whose
precondition is ``on(A,B),clear(A),clear(C)'', whose add is ``on(A,C),
clear(B)'' and whose del is ``on(A,B), clear(C)''. A 2-split is given
by ``A,B'' and ``A,C''. We get the actions ``move1(A,B)'' with pre
``processing-none, on(A,B), clear(A)'', add ``do-move2, clear(B),
arg1(A), arg2(B)'', del ``on(A B), processing-none''; and
``move2(A,C)''with pre ``do-move2, clear(C), arg1(A)'', add ``on(A C),
processing-none, arg3(C)'', del ``clear(C), do-move2''.

@@

>>>>

For example, consider the blocksworld operator ``move(A,B,C)'', whose
precondition is ``on(A,B),clear(A),clear(C)'', whose add is ``on(A,C),
clear(B)'' and whose del is ``on(A,B), clear(C)''. A 2-split is given
by ``A,B'' and ``A,C''. We get the actions ``move1(A,B)'' with pre
``processing-none, on(A,B), clear(A)'', add ``do-move2, clear(B),
arg1(A), arg2(B)'', del ``on(A B), processing-none''; and
``move2(A,C)'' with pre ``do-move2, clear(C), arg1(A)'', add ``on(A C),
processing-none, arg3(C)'', del ``clear(C), do-move2''.

@@

One important remark concerning planner performance: splitting up
operators in the above way may produce dead-ends. After the first
operator is applied, we do not have $processing\mbox{-}none$ anymore
and thus there is no chance of reaching the goal other than by
completing the split, i.e., reaching an operator
$do\mbox{-}o_k$. That, however, may not be possible. As of the first
operator, we commit to a partial variable instantiation $i$ for the
operator in question, in terms of the $arg_1(.), \dots, arg_l(.)$
predicates. Any completion of the split -- at any point during the
execution of the split -- must stick to $i$. It may be the case that
some ``future'' precondition, i.e., a precondition included in a
later-on part of the split, cannot be satisfied in the current state
in conformance with $i$. As a simple example, if we enrich the
blocksworld with an additional static predicate ``can-stack(x y)'' and
allow ``move(A,B,C)'' only if ``can-stack(A C)'' holds, then in the
above split ``move2(A,C)'' has the additional precondition
``can-stack(A C)''. However, there may not exist ``C'' for which both
``can-stack(A C)'' and ``clear(C)'' hold. In this case, upon
application of ``move1(A,B)'' we're in a dead-end. However, if the
planner is using a relaxed-plan based heuristic, then these dead-ends
``do not do a lot of damage'', in the sense that for any such dead-end
state $s$, no relaxed plan will exist and thus the heuristic value
will be $\infty$ and the state will be immediately pruned. The simple
reason for this is that $processing\mbox{-}none$ must be re-achieved
also in the relaxed plan, and doing so is not possible because at some
point along the split sequence no more actions will be applicable.

### 2011-05-23 tags/crisp-1.1.6/trunk/Papers/operator-splitting/total.tex
[maven-release-plugin]  copy for tag crisp-1.1.6

We herein begin to address this question, focussing on one particular
technical detail that affects the pre-processing (and in certain ways
also the search techniques) of many existing planners: the number of
operator parameters. For example, in the development of the Pipesworld
domain as used at IPC 2004, one of the essential steps was to split up
some of the operators, so that each part would have less parameters
and thus be more feasible for common pre-processing techniques. We
herein provide an automatic method that does just that. We evaluate
the method on IPC benchmarks, as well as on the domain of sentence
generation, where pre-processing, in particular related to the number
of operator parameters, is known to be one of the major obstacles.
\end{abstract}

>>>>

We herein begin to address this question, focussing on one particular
technical detail that affects the pre-processing (and in certain ways
also the search techniques) of many existing planners: the number of
operator parameters. For example, in the development of the Pipesworld
domain as used at IPC 2004, one of the essential steps was to split up
some of the operators, so that each part would have less parameters
and thus be more feasible for common pre-processing techniques. We
herein provide an automatic method that does just that. We evaluate
the method on IPC benchmarks, as well as on the domain of sentence
generation, where pre-processing, in particular related to the number
of operator parameters, is known to be one of the major obstacles.

\joerg{NOTE: instead of selling this as a re-writing technique, we
  could just as well sell it is a way to ``get complex operators
  through common pre-processors'': after all, once the pre-processing
  is through, inside the planner we could easily ``unsplit'' the
  operators again, yielding exactly the same grounded representation
  that we would have got from the original operators. ... the
  advantage is that we will not suffer from any potential search
  degradation, and can focus exclusively on reducing instantiation
  time. the disadvantage of course is in cases where the grounded
  representation is prohibitively big -- the number of ``original''
  ground ops can be exponentially (in the number of op params) larger
  than the number of ``splitted'' ground ops.}
\end{abstract}

### 2011-05-23 tags/crisp-1.1.6/trunk/Papers/pcrisp-10/crisp.tex
[maven-release-plugin]  copy for tag crisp-1.1.6

\section{Sentence Generation as Planning}
\label{sec:crisp}
In this section we review the original non-statistical {\sc crisp} system \cite{kollerstone2007}. 
Following {\sc spud} \cite{Stone2003a}, {\sc crisp} is based on a declarative description of the sentence generation problem using {\sc tag}. Given a knowledge base, a communicative goal and a grammar, we require to find a grammatical {\sc tag} derivation that is consistent with this knowledge base and satisfies a communicative goal. A number of semantic and pragmatic constraints that must be satisfied by the solution can be added, for instance to enforce generation of unambiguous referring expressions. \newcite{kollerstone2007} describe how to encode this problem into an AI planning problem which can be solved efficiently by off-the-shelf planners. We describe the general mechanism in the following section and then review the encoding into planning in section \ref{ssec:crispdomain}. 
\begin{figure*}[th]
\begin{center}
\includegraphics[width=.8\textwidth]{figures/grammar.pdf}
\caption{\label{fig:grammar}(a) An example grammar with semantic content. (b) A derivation for ``the cat eats the raw fish''. }
\end{center}
\end{figure*}

\subsection{Sentence Generation in CRISP}
Like {\sc spud}, {\sc crisp} uses an {\sc ltag} in which elementary trees are assigned semantic content. Each node in a {\sc crisp} elementary tree is associated with a semantic role. Semantic content is expressed as a set of literals, encoding relations between these roles. All nodes that dominate the lexical anchor are assigned the role `self', which intuitively corresponds to the event or individual described by this tree. Fig.~\ref{fig:grammar}(a) shows an example grammar of this type.

In a derivation we may only include elementary trees whose semantic content has an instantiation in the knowledge base. For each substitution and adjunction, the semantic role associated with the role of the target node is unified with the `self' role of the child tree. For example, given the knowledge base $\{\f{cat}(c), \f{fish}(f_1), \f{raw}(f_1), \f{fish}(f_2), \f{eats}(e,c,f_1)\}$ and the grammar in \ref{fig:grammar}(a), {\sc crisp} could produce the derivation in \ref{fig:grammar}(b). Notice that {\sc crisp} can generate the unambiguous referring expression `the raw fish' for $f_1$, to distinguish it from $f_2$.

\subsection{CRISP Planning Domains} 
\label{ssec:crispdomain}
Before we describe {\sc crisp}'s encoding of sentence generation as planning, we briefly review AI planning in general.  A planning state is a conjunction of first order literals describing relations between some individuals. A planning problem consist of an initial state, a set of goal states and a set of planning operators that describe possible state transitions.  A \emph{plan} is any sequence of actions (instantiated operators) that leads from the initial state to one of the goal states. Planning problems can be solved efficiently by general purpose planning systems such as {\sc ff} \cite{hoffmannnebel2001}.

In {\sc crisp}, planning states correspond to partial {\sc tag} derivations and record open substitution and adjunction sites, semantic individuals associated with them, and parts of the communicative goal that have not yet been expressed.  The initial state also encodes the knowledge base and the communicative goal. Each planning operator contributes a new elementary tree to the derivation and at the same time can satisfy part of the communicative goal, as described in the previous section. In a goal state there are no open substitution sites left and all literals in the communicative goal have been expressed. 
\begin{figure}[t]
\begin{center}
\planaction{\bf subst-t28-eats-S(u,~x1,~x2,~x3)}{referent(u,~x1),\\ subst(S,~u), eats(x1,~x2,~x3)}{
$\lnot$needtoexpr(pred-eats,~x1,~x2,~x3),\\ $\lnot$subst(S,~u),\\
subst(NP,~subj), subst(NP,~obj),\\ referent(subj,~x2), referent(obj,~x3)\\
adj(VP, u), adj(V, u), adj(S, u)}\\\smallskip

@@

\planaction{\bf adj-t5-raw-N(u,~x1)}{referent(u,~x1),\\ adj(N,~u), raw(x1)}{
$\lnot$needtoexpr(pred-raw,~x1),\\ $\lnot$adj(N,~u)}\\\smallskip
\end{center}
\caption{\label{fig:crisp-operators} {\sc crisp} operators for some of the elementary trees in Fig.~\ref{fig:grammar}.}
\end{figure}

@@

\newcite{KolHof10} report on experiments that show CRISP can generate sentences with the large-scale XTAG grammar \cite{xtag2001} quite efficiently.  However, because CRISP has no notion of how ``good'' a generated sentence is compared to other grammatical alternatives, it will sometimes compute dispreferred sentences with large grammars.  This is especially true for treebank-induced grammars, which tend to overgenerate and rely on statistical methods to rank good sentences highly.  Fig.~\ref{fig:overgen} illustrates this problem.  Assuming a (treebank) grammar that includes trees for both right adjoining (t13) and left adjoining PPs (t252), both derivations (a) and (b) are {\it grammatical} derivations that satisfy the same communicative goal. However, most readers disprefer the reading in (b).  Clearly, to use {\sc crisp} with such a grammar we need a method of distinguishing good derivations from bad ones.

>>>>

\section{Sentence Generation as Planning}
\label{sec:crisp}
In this section we review the original non-statistical CRISP system \cite{kollerstone2007}. 
Following SPUD \cite{Stone2003a}, CRISP is based on a declarative description of the sentence generation problem using TAG. Given a knowledge base, a communicative goal and a grammar, we require to find a grammatical TAG derivation that is consistent with this knowledge base and satisfies a communicative goal. A number of semantic and pragmatic constraints that must be satisfied by the solution can be added, for instance to enforce generation of unambiguous referring expressions. \newcite{kollerstone2007} describe how to encode this problem into an AI planning problem which can be solved efficiently by off-the-shelf planners. We describe the general mechanism in the following section and then review the encoding into planning in section \ref{ssec:crispdomain}. 
\begin{figure*}[th]
\begin{center}
\includegraphics[width=.8\textwidth]{figures/grammar.pdf}
\caption{\label{fig:grammar}(a) An example grammar with semantic content. (b) A derivation for ``the cat eats the raw fish''. }
\end{center}
\end{figure*}

\subsection{Sentence Generation in CRISP}
Like SPUD, CRISP uses an LTAG in which elementary trees are assigned semantic content. Each node in a CRISP elementary tree is associated with a semantic role. Semantic content is expressed as a set of literals, encoding relations between these roles. All nodes that dominate the lexical anchor are assigned the role `self', which intuitively corresponds to the event or individual described by this tree. Fig.~\ref{fig:grammar}(a) shows an example grammar of this type.

In a derivation we may only include elementary trees whose semantic content has an instantiation in the knowledge base. For each substitution and adjunction, the semantic role associated with the role of the target node is unified with the `self' role of the child tree. For example, given the knowledge base $\{\f{cat}(c), \f{fish}(f_1), \f{raw}(f_1), \f{fish}(f_2), \f{eats}(e,c,f_1)\}$ and the grammar in \ref{fig:grammar}(a), CRISP could produce the derivation in \ref{fig:grammar}(b). Notice that CRISP can generate the unambiguous referring expression `the raw fish' for $f_1$, to distinguish it from $f_2$.

\subsection{CRISP Planning Domains} 
\label{ssec:crispdomain}
Before we describe CRISP's encoding of sentence generation as planning, we briefly review AI planning in general.  A planning state is a conjunction of first order literals describing relations between some individuals. A planning problem consist of an initial state, a set of goal states and a set of planning operators that describe possible state transitions.  A \emph{plan} is any sequence of actions (instantiated operators) that leads from the initial state to one of the goal states. Planning problems can be solved efficiently by general purpose planning systems such as {\sc ff} \cite{hoffmannnebel2001}.

In CRISP, planning states correspond to partial TAG derivations and record open substitution and adjunction sites, semantic individuals associated with them, and parts of the communicative goal that have not yet been expressed.  The initial state also encodes the knowledge base and the communicative goal. Each planning operator contributes a new elementary tree to the derivation and at the same time can satisfy part of the communicative goal, as described in the previous section. In a goal state there are no open substitution sites left and all literals in the communicative goal have been expressed. 
\begin{figure}[t]
\begin{center}
\planaction{\bf subst-t28-eats-S(u,~x1,~x2,~x3)}{referent(u,~x1),\\ subst(S,~u), eats(x1,~x2,~x3)}{
$\lnot$needtoexpr(pred-eats,~x1,~x2,~x3),\\ $\lnot$subst(S,~u),\\
subst(NP,~subj), subst(NP,~obj),\\ referent(subj,~x2), referent(obj,~x3)\\
adj(VP, u), adj(V, u), adj(S, u)}\\\smallskip

@@

\planaction{\bf adj-t5-raw-N(u,~x1)}{referent(u,~x1),\\ adj(N,~u), raw(x1)}{
$\lnot$needtoexpr(pred-raw,~x1),\\ $\lnot$adj(N,~u)}\\\smallskip
\end{center}
\caption{\label{fig:crisp-operators} CRISP operators for some of the elementary trees in Fig.~\ref{fig:grammar}.}
\end{figure}

@@

\newcite{KolHof10} report on experiments that show CRISP can generate sentences with the large-scale XTAG grammar \cite{xtag2001} quite efficiently.  However, because CRISP has no notion of how ``good'' a generated sentence is compared to other grammatical alternatives, it will sometimes compute dispreferred sentences with large grammars.  This is especially true for treebank-induced grammars, which tend to overgenerate and rely on statistical methods to rank good sentences highly.  Fig.~\ref{fig:overgen} illustrates this problem.  Assuming a (treebank) grammar that includes trees for both right adjoining (t13) and left adjoining PPs (t252), both derivations (a) and (b) are {\it grammatical} derivations that satisfy the same communicative goal. However, most readers disprefer the reading in (b).  Clearly, to use CRISP with such a grammar we need a method of distinguishing good derivations from bad ones.

### 2011-05-23 tags/crisp-1.1.6/trunk/Papers/pcrisp-10/experiments.tex
[maven-release-plugin]  copy for tag crisp-1.1.6

We now report on some experiments with \pcrisp. We are interested in the output quality and runtime behavior of different combinations of planning systems and probability models. Like \crisp, \pcrisp\ is an integrated sentence generation system capable of generating referring expressions during surface realization.  However, for lack of an appropriate evaluation dataset for full sentence generation, we only evaluate \pcrisp\ on a realization task here.  Further experiments and more details can be found in \newcite{bauer2009}.

\subsection{Evaluation Data}
For our experiments we use an \ltag\ grammar and treebank that was automatically extracted from the Wall Street Journal using the algorithm described by \newcite{chenschanker2004}.

This algorithm outputs a grammar that allows multiple adjunctions at the same node. For such a grammar {\sc ptag} is not a suitable probability model. Models that can deal with multiple adjunction are discussed by \newcite{nasrrambow2006}, but employing them would require non-trivial modifications to our encoding as a metric planning problem. We therefore preprocess the treebank by linearizing multiple adjunctions. Furthermore, we reattach prepositions to the trees they substitute in, to increase the expressiveness of the bilexical probability distribution.

@@

We estimate our probability models on section 1 to 23 of the converted WSJ using maximum likelihood estimation. We use section 0 as a testing set. However, since the number of \pcrisp\ operators grows quadratically with the grammar size, generating long sentences requires too much time to run batch experiments.  We therefore restrict our evaluation to the 416 sentences in Section 0 that are shorter than 16 words.

@@

\subsection{Generation tree accuracy}
\label{ssec:eval-measures}
To evaluate the output quality of our statistical generation system we compare the system output $O$ against the reference sentence $R$ in the treebank from which the input representation was generated. We adopt the {\bf generation tree accuracy} (GTA) measure discussed by \cite{bangalorerambowwhittaker2001}.
This measure is computed by first creating a list of all 'treelets' from the reference derivation tree $D$. A `treelet' is a subtree consisting only of a node and its direct children. For each treelet we calculate the edit distance, sum the distances over all treelets and then divide by the total length of the reference string:
$$1 - \frac{\sum\limits_{d \in treelets(D)} editDist(O|_d, R|_d)}{\textit{length}(R)},$$
where $D$ is the reference derivation tree and $S|_d$ are the tokens associated with the nodes of treelet $d$ in the order they appear in $S$ (if at all).  Edit distance is modified to treat insertion-deletion pairs as single movement errors.  Compared to a purely string-based metric like {\sc bleu}, GTA penalizes swapped words less harshly if they can be explained by local tree movements.

@@

Table \ref{results} presents the results of the experiment for five different generation systems.  We compare three variants of \pcrisp: the fully lexicalized PTAG model (``PTAG''), the fully unlexicalized model (``unlexicalized''), and a linear interpolation model (``interpolation'') in which we (manually) set the weight of level 2 to 0.9 and the weight of level 3 to 0.  We also list results for the non-statistical \crisp\ system of  \newcite{kollerstone2007} (``\crisp'') and the greedy search heuristic used by \spud.  All these systems are based on a reimplementation of the FF planner \cite{hoffmannnebel2001}, to which we added a search heuristic that takes action costs into account for the \pcrisp\ systems.

@@

The table confirms that the non-statistical \crisp\ system has considerable trouble reconstructing the original sentences from the treebank, with a mean GTA of 0.66.  This is still better than our reimplementation of \spud, which fails to recover from early mistakes of its greedy search heuristic for every single sentence in the test set.

By contrast, the fully lexicalized \pcrisp\ model achieves a much better mean GTA.  However, this comes at the cost of a very low success rate of only 10\%, reflecting a serious data sparseness problem on unseen inputs.  The data sparseness problem is reduced in the unlexicalized version of \pcrisp, but this comes at the cost of decreased accuracy and much increased runtimes.  The linear interpolation model strikes a balance between these two, by improving the success rate over the lexicalized model while sacrificing only a small amount of accuracy.  This suggests that smoothing is a promising approach to balancing coverage, efficiency, and accuracy, but clearly further experimentation is needed to substantiate this.

\begin{table}
    \begin{center}
    \begin{tabular}{|l||l||r|r|r|}
    \hline
     {\bf System }  & {\bf gta}& {\bf success} & {\bf fail} & {\bf timeout} \\ \hline 
     \spud         & n/a & 0\% & 100\% & 0\%  \\ \hline 
     \crisp         &  0.66 & 45\% & 42\% & 13\% \\ \hline
     PTAG &  0.90 &  10\% &88\% & 2\% \\ \hline
     unlexicalized & 0.74 & 62\% & 16\% & 22\%\\ \hline
     interpolation & 0.88 & 19\% & 74\% & 7\% \\ \hline
    \end{tabular}
    \end{center}
    \caption{\label{results} Results for the realization experiment.} 
\end{table}

>>>>

We now report on some experiments with PCRISP. We are interested in the output quality and runtime behavior of different combinations of planning systems and probability models. Like CRISP, PCRISP is an integrated sentence generation system capable of generating referring expressions during surface realization.  However, for lack of an appropriate evaluation dataset for full sentence generation, we only evaluate PCRISP on a realization task here.  Further experiments and more details can be found in \newcite{bauer2009}.

\subsection{Evaluation Data}
For our experiments we use an LTAG grammar and treebank that was automatically extracted from the Wall Street Journal using the algorithm described by \newcite{chenschanker2004}.

This algorithm outputs a grammar that allows multiple adjunctions at the same node. For such a grammar PTAG is not a suitable probability model. Models that can deal with multiple adjunction are discussed by \newcite{nasrrambow2006}, but employing them would require non-trivial modifications to our encoding as a metric planning problem. We therefore preprocess the treebank by linearizing multiple adjunctions. Furthermore, we reattach prepositions to the trees they substitute in, to increase the expressiveness of the bilexical probability distribution.

@@

We estimate our probability models on section 1 to 23 of the converted WSJ using maximum likelihood estimation. We use section 0 as a testing set. However, since the number of PCRISP operators grows quadratically with the grammar size, generating long sentences requires too much time to run batch experiments.  We therefore restrict our evaluation to the 416 sentences in Section 0 that are shorter than 16 words.

@@

\subsection{Generation tree accuracy}
\label{ssec:eval-measures}
To evaluate the output quality of our statistical generation system we compare the system output $O$ against the reference sentence $R$ in the treebank from which the input representation was generated. We adopt the {\bf generation tree accuracy} (GTA) measure discussed by \cite{bangalorerambowwhittaker2001}.
This measure is computed by first creating a list of all 'treelets' from the reference derivation tree $D$. A `treelet' is a subtree consisting only of a node and its direct children. For each treelet we calculate the edit distance, sum the distances over all treelets and then divide by the total length of the reference string:
$$1 - \frac{\sum\limits_{d \in treelets(D)} editDist(O|_d, R|_d)}{\textit{length}(R)},$$
where $D$ is the reference derivation tree and $S|_d$ are the tokens associated with the nodes of treelet $d$ in the order they appear in $S$ (if at all).  Edit distance is modified to treat insertion-deletion pairs as single movement errors.  Compared to a purely string-based metric like BLEU, GTA penalizes swapped words less harshly if they can be explained by local tree movements.

@@

Table \ref{results} presents the results of the experiment for five different generation systems.  We compare three variants of PCRISP: the fully lexicalized PTAG model (``PTAG''), the fully unlexicalized model (``unlexicalized''), and a linear interpolation model (``interpolation'') in which we (manually) set the weight of level 2 to 0.9 and the weight of level 3 to 0.  We also list results for the non-statistical CRISP system of  \newcite{kollerstone2007} (``CRISP'') and the greedy search heuristic used by SPUD.  All these systems are based on a reimplementation of the FF planner \cite{hoffmannnebel2001}, to which we added a search heuristic that takes action costs into account for the PCRISP systems.

@@

The table confirms that the non-statistical CRISP system has considerable trouble reconstructing the original sentences from the treebank, with a mean GTA of 0.66.  This is still better than our reimplementation of SPUD, which fails to recover from early mistakes of its greedy search heuristic for every single sentence in the test set.

By contrast, the fully lexicalized PCRISP model achieves a much better mean GTA.  However, this comes at the cost of a very low success rate of only 10\%, reflecting a serious data sparseness problem on unseen inputs.  The data sparseness problem is reduced in the unlexicalized version of PCRISP but this comes at the cost of decreased accuracy and much increased runtimes.  The linear interpolation model strikes a balance between these two, by improving the success rate over the lexicalized model while sacrificing only a small amount of accuracy.  This suggests that smoothing is a promising approach to balancing coverage, efficiency, and accuracy, but clearly further experimentation is needed to substantiate this.

\begin{table}
    \begin{center}
    \begin{tabular}{|l||l||r|r|r|}
    \hline
     {\bf System }  & {\bf gta}& {\bf success} & {\bf fail} & {\bf timeout} \\ \hline 
     SPUD         & n/a & 0\% & 100\% & 0\%  \\ \hline 
     CRISP         &  0.66 & 45\% & 42\% & 13\% \\ \hline
     PTAG &  0.90 &  10\% &88\% & 2\% \\ \hline
     unlexicalized & 0.74 & 62\% & 16\% & 22\%\\ \hline
     interpolation & 0.88 & 19\% & 74\% & 7\% \\ \hline
    \end{tabular}
    \end{center}
    \caption{\label{results} Results for the realization experiment.} 
\end{table}

### 2011-05-23 tags/crisp-1.1.6/trunk/Papers/pcrisp-10/related.tex
[maven-release-plugin]  copy for tag crisp-1.1.6

Statistical methods are popular for surface realization, but have not been used in systems that integrate sentence planning. Most statistical generation approaches follow a generate-and-select strategy, first proposed by \newcite{knight1995} in their {\sc nitrogen} system. Such systems generate a set of candidate sentences using a (possibly overgenerating) grammar and then select the best output sentence by applying a statistical language model. This family includes systems such as {\sc HAL}ogen \cite{langkildeknight1998,langkilde2000} and OpenCCG \cite{whitebaldridge2003}.  The {\sc Fergus} system \cite{bangalorerambow2000} is a variant of this approach which, like {\sc pcrisp}, employs {\sc ltag}. It first assigns elementary trees to each entity in the input sentence plan using a statistical tree model and then computes the most likely derivation using only these trees with an $n$-gram model on the output sentence. An alternative to the $n$-gram based generate and select approach is to use a probabilistic grammar model, like {\sc   ptag}, trained on automatic parses \cite{zhongstent2005}. A related approach uses a model over local decisions of the generation system itself \cite{belz2008}. Both models can either be used to discriminate a set of output candidates, or more directly to choose the next best decision locally. Our approach is similar in that it uses {\sc ptag} to find the most likely output structure. However, the previous work discussed so far addresses surface realization only. We extend this to a statistical NLG algorithm which does surface realization and sentence planning at the same time.

Our treatment of integrated sentence planning and surface realization as planning is inherited directly from {\sc crisp} \cite{kollerstone2007}.  Planning has long played a role in generation, but has focused on discourse planning instead of specifically addressing sentence generation \cite{hovy1988,appelt1992}. The applicability of these ideas was limited at that time because efficient planning technology was not available. Recently the development of more efficient planning algorithms \cite{hoffmannnebel2001} spawned a renewed interest in planning for NLG.  {\sc Crisp} uses such algorithms to efficiently solve the sentence generation problem defined by {\sc spud} \cite{Stone2003a}. {\sc Spud}, which instead uses an incomplete greedy algorithm, is based on an {\sc ltag} whose trees are augmented with semantic and pragmatic constraints. Given a communicative goal, a solution to the {\sc spud} problem realizes this goal and simultaneously selects referring expressions. The next section explores {\sc crisp} in more detail.

>>>>

Statistical methods are popular for surface realization, but have not been used in systems that integrate sentence planning. Most statistical generation approaches follow a generate-and-select strategy, first proposed by \newcite{knight1995} in their NITROGEN system. Such systems generate a set of candidate sentences using a (possibly overgenerating) grammar and then select the best output sentence by applying a statistical language model. This family includes systems such as HALogen \cite{langkildeknight1998,langkilde2000} and OpenCCG \cite{whitebaldridge2003}.  The FERGUS system \cite{bangalorerambow2000} is a variant of this approach which, like PCRISP, employs TAG. It first assigns elementary trees to each entity in the input sentence plan using a statistical tree model and then computes the most likely derivation using only these trees with an n-gram model on the output sentence. An alternative to the n-gram based generate and select approach is to use a probabilistic grammar model, like PTAG, trained on automatic parses \cite{zhongstent2005}. A related approach uses a model over local decisions of the generation system itself \cite{belz2008}. Both models can either be used to discriminate a set of output candidates, or more directly to choose the next best decision locally. Our approach is similar in that it uses PTAG to find the most likely output structure. However, the previous work discussed so far addresses surface realization only. We extend this to a statistical NLG algorithm which does surface realization and sentence planning at the same time.

Our treatment of integrated sentence planning and surface realization as planning is inherited directly from CRISP \cite{kollerstone2007}.  Planning has long played a role in generation, but has focused on discourse planning instead of specifically addressing sentence generation \cite{hovy1988,appelt1992}. The applicability of these ideas was limited at that time because efficient planning technology was not available. Recently the development of more efficient planning algorithms \cite{hoffmannnebel2001} spawned a renewed interest in planning for NLG. CRISP uses such algorithms to efficiently solve the sentence generation problem defined by SPUD \cite{Stone2003a}. SPUD, which instead uses an incomplete greedy algorithm, is based on a TAG  whose trees are augmented with semantic and pragmatic constraints. Given a communicative goal, a solution to the SPUD problem realizes this goal and simultaneously selects referring expressions. The next section explores CRISP in more detail.

### 2011-05-23 tags/crisp-1.1.6/trunk/Papers/reasoning-challenges/abstract.tex
[maven-release-plugin]  copy for tag crisp-1.1.6

\begin{abstract}
We revisit the structure of the International Planning Competition (IPC)
and propose a new ``rolling'' competition, as a permanent fixture for the
planning community. A central competition server would continually evaluate
planners and generate performance data, while members of the community
could directly upload new planners and problem domains, with the server
automatically reevaluating existing planners as necessary. Real-time
results would continuously be available, while more formal competitions
could be ``run'' at regular intervals by producing snapshots of current
planners and problems. Researchers would have access to a central
repository with the latest domains and performance statistics, running on a
common hardware platform, plus a facility for related activities such as
posting challenges and demonstrating new technologies. Since much of the
proposed infrastructure is not tied to a specific research area, this
approach offers the possibility of joint ventures with other research
communities.
\end{abstract}

>>>>

\begin{abstract}
We revisit the structure of the International Planning Competition (IPC)
and propose a new ``rolling'' competition, as a permanent fixture for the
planning community. A central competition server would continually evaluate
planners and generate performance data, while members of the community
could directly upload new planners and problem domains, with the server
automatically reevaluating existing planners as necessary. Real-time
results would continuously be available, while more formal competitions
could be ``run'' at regular intervals by producing snapshots of current
planners and problems. Researchers would have access to a central
repository with the latest domains and performance data, running on a
common hardware platform, plus a social networking forum for posting
challenges, announcing new technologies, and discussion. Since much of
the proposed infrastructure is not tied to a specific research area, this
approach offers the possibility of joint ventures with other research
communities.
\end{abstract}
